---
title: "Simple Model Exploration 2"
author: "gp"
date: "5/6/2020"
output: html_document
---

###CLEANING DATA###


Libraries + Data
```{r}
library(tidyverse)
library(readxl)
data<-simple_cleaning_506%>%
  select(drug_x, pmid, predicted, detrimental_probability, effective_probability, inconclusive_probability, noeffect_probability, disease, disease_cleaned, studytype1,studytype2, studytype3, studytype4, studytype5, studytype6, studytype7)
```

What are the unique studytypes listed in the data for studytype1?
```{r}
unique(data$studytype1)
```
Study Types1 Mapped to Terms in CWR4C Workflow: 

journal article---------------No MAP
comparative study-------------clinical trial
controlled clinical trial-----clinical trial
letter------------------------No MAP
case reports-------------------clinical case report
clinical trial----------------clinical trial
english abstract--------------NO MAP
evaluation study--------------No MAP

CWR4C Studytypes:

meta-analysis
clinical trial
clinical observational
meta-synthesis
clinical case report
preclinical in vivo
preclinical in vitro
other

Mapping studytypes1 in data with a match to CWR4C Studytypes: 
```{r}
for(x in 1: count(data)$n){
  if (data$studytype1[x]=="comparative study")
    {data$studytype1[x]= "clinical trial"}
  if (data$studytype1[x]=="controlled clinical trial")
    {data$studytype1[x]= "clinical trial"}
  if (data$studytype1[x]=="case reports")
    {data$studytype1[x]="clinical case report"}
}
```

Finding all studies in the data with mapped terms for studytype1:
```{r}
mapped<-data%>%
  filter(studytype1=="clinical trial"|studytype1=="clinical case report")
```

Finding all studies in the data without mapped terms for studytype1
```{r}
not_mapped<-data%>%
  filter(studytype1!="clinical trial")%>%
  filter(studytype1!="clinical case report")
unique(not_mapped$studytype1)
```

Assumption: 
changing studies where studytye 1 = "journal article" (and it has no other label in studytype2) to "preclinical in vitro"
```{r}
journal_nolabel<-not_mapped%>%
  filter(studytype1=="journal article"& is.na(studytype2))

for(x in 1: count(journal_nolabel)$n){
  if (journal_nolabel$studytype1[x]=="journal article")
  {journal_nolabel$studytype1[x]= "preclinical in vitro"}
}
```

combining mapped and journal_nolabel subdatasets()
```{r}
final_data<-rbind(mapped, journal_nolabel)
```

Finding studies labled as a journal article with a second label(studytype2 not empty)
Seeing the unique studytypes in studytype2
```{r}
journal_studytype2<-not_mapped%>%filter(studytype1=="journal article"&!is.na(studytype2))

unique(journal_studytype2$studytype2)
```
Journal-Study Types2 Mapped to Terms in CWR4C Workflow: 

research support--------------------No MAP
randomized controlled trial---------clinical trial
observational study-----------------clinical observational
multicenter study-------------------clinical trial


CWR4C Studytypes:

meta-analysis
clinical trial
clinical observational
meta-synthesis
clinical case report
preclinical in vivo
preclinical in vitro
other

Mapping studytypes1 in data with a match to CWR4C Studytypes:
```{r}
for(x in 1:count(journal_studytype2)$n){
  if (journal_studytype2$studytype2[x]=="randomized controlled trial")
    {journal_studytype2$studytype1[x]= "clinical trial"}
  if (journal_studytype2$studytype2[x]=="observational study")
    {journal_studytype2$studytype1[x]= "clinical observational"}
  if (journal_studytype2$studytype2[x]=="multicenter study")
    {journal_studytype2$studytype1[x]=="clinical trial"}
}
```
Finding all studies in the data with mapped terms for studytype2 onto studytype 1 column:

```{r}
mapped2<-journal_studytype2%>%
  filter(studytype2=="clinical trial"|studytype1=="clinical observational")
```

combining mapped2 articles to final_data
```{r}
rbind(final_data,mapped2)
```




Finding all studies in the data without mapped terms for studytype2
```{r}
not_mapped2<-journal_studytype2%>%
  filter(studytype2!="randomized controlled trial")%>%
  filter(studytype2!="observational study")%>%
  filter(studytype2!="multicenter study")
unique(not_mapped2$studytype2)
```
Finding the studies where studytype1=journal article, studytype2 is research support and studytype 3 is NOT empty
```{r}
journal_research_studytype3<-not_mapped2%>%filter(!is.na(studytype3))
unique(journal_research_studytype3$studytype4)
```

```{r}

letter<-not_mapped%>%filter(studytype1=="letter")
english_abstract<-not_mapped%>%filter(studytype1=="english abstract")
evaluation_study<-not_mapped%>%filter(studytype1=="evaluation study")

other<-rbind(letter,english_abstract,evaluation_study)
other
```
manually searched through all studytype columns for the "other" not mapped studytypes for studytype1(letter, english abstract, evaluation study)-only the one study as a letter needed to be mapped:
```{r}
map_letter<-data%>%filter(studytype1=="letter")%>%mutate(studytype1="clinical trial")
map_letter
```
Add map_letter to final_data
```{r}
rbind(map_letter, final_data)
```
final_data is smaller than original data because studies that were labled as journal article-research support-NA were ignored 

###SCORING MODEL###

```{r}
final_data<-final_data%>%mutate(weight=0)
num_studies<-final_data%>%count()

for(x in 1:num_studies$n){
  if((final_data$studytype1[x] == "meta-analysis") & (final_data$predicted[x]=="Effective")){final_data$weight[x]=8}
  if((final_data$studytype1[x]== "meta-analysis") & (final_data$predicted[x]=="Detrimental")){final_data$weight[x]=-8}
  if((final_data$studytype1[x]== "meta-analysis") & (final_data$predicted[x]=="Inconclusive")){final_data$weight[x]=-4}
  
  if((final_data$studytype1[x]== "clinical trial") & (final_data$predicted[x]=="Effective")){final_data$weight[x]=7}
  if((final_data$studytype1[x]== "clinical trial") & (final_data$predicted[x]=="Detrimental")){final_data$weight[x]=-7}
  if((final_data$studytype1[x]== "clinical trial") & (final_data$predicted[x]=="Inconclusive")){final_data$weight[x]=-3.5}
  
  if((final_data$studytype1[x]== "clinical observational") & (final_data$predicted[x]=="Effective")){final_data$weight[x]=6}
  if((final_data$studytype1[x]== "clinical observational") & (final_data$predicted[x]=="Detrimental")){final_data$weight[x]=-6}
  if((final_data$studytype1[x]== "clinical observational") & (final_data$predicted[x]=="Inconclusive")){final_data$weight[x]=-3}
  
  if((final_data$studytype1[x]== "meta-synthesis") & (final_data$predicted[x]=="Effective")){final_data$weight[x]=5}
  if((final_data$studytype1[x]== "meta-synthesis") & (final_data$predicted[x]=="Detrimental")){final_data$weight[x]=-5}
  if((final_data$studytype1[x]== "meta-synthesis") & (final_data$predicted[x]=="Inconclusive")){final_data$weight[x]=-2.5}

  if((final_data$studytype1[x]== "clinical case report") & (final_data$predicted[x]=="Effective")){final_data$weight[x]=4}
  if((final_data$studytype1[x]== "clinical case report") & (final_data$predicted[x]=="Detrimental")){final_data$weight[x]=-4}
  if((final_data$studytype1[x]== "clinical case report") & (final_data$predicted[x]=="Inconclusive")){final_data$weight[x]=-2} 
  
  if((final_data$studytype1[x]== "preclinical in vivo") & (final_data$predicted[x]=="Effective")){final_data$weight[x]=3}
  if((final_data$studytype1[x]== "preclinical in vivo") & (final_data$predicted[x]=="Detrimental")){final_data$weight[x]=-3}
  if((final_data$studytype1[x]== "preclinical in vivo") & (final_data$predicted[x]=="Inconclusive")){final_data$weight[x]=-1.5}
  
  if((final_data$studytype1[x]== "preclinical in vitro") & (final_data$predicted[x]=="Effective")){final_data$weight[x]=2}
  if((final_data$studytype1[x]== "preclinical in vitro") & (final_data$predicted[x]=="Detrimental")){final_data$weight[x]=-2}
  if((final_data$studytype1[x]== "preclinical in vitro") & (final_data$predicted[x]=="Inconclusive")){final_data$weight[x]=-1}
}

scores<-final_data%>%group_by(drug_x,disease_cleaned)%>%
        summarize(score=sum(weight),count=n())%>%
        arrange(desc(score))
scores

```
***SCORING MODEL DRUG ONLY***
```{r}
final_data<-final_data%>%mutate(weight=0)

for(x in 1:num_studies$n){
  if(final_data$predicted[x] == "Effective"){final_data$weight[x]=10}
  if(final_data$predicted[x] == "Detrimental"){final_data$weight[x]=-10}
  if(final_data$predicted[x] == "Inconclusive"){final_data$weight[x]=-5}
}

scores<-final_data%>%group_by(drug_x)%>%
        summarize(score=sum(weight))%>%
        arrange(desc(score))
scores
  
```
Number of Studies per drug

```{r}
num_studies_drug<-final_data%>%group_by(drug_x)%>%count()
num_studies_drug
```
LOOKING AT STUDIES LABELED "INCONCLUSIVE" 
```{r}
final_data<-final_data%>%mutate(ie=0,id=0,ine=0,min_difference=0, check=0)

inconclusive<-final_data%>%filter(predicted =="Inconclusive")%>%
  mutate(ie=inconclusive_probability-effective_probability, 
         id = inconclusive_probability-detrimental_probability,
         ine= inconclusive_probability-noeffect_probability)
inconclusive<-inconclusive%>%
  mutate(min_difference=apply(X=inconclusive[18:20], MARGIN=1, FUN=min))%>%
  mutate(check=min_difference==ie)

check<-inconclusive%>%filter(check==FALSE)

not_inconclusive<-final_data%>%filter(predicted!="Inconclusive")
changed_inconclusive<-inconclusive%>%filter(min_difference<=0.15)%>%
  mutate(predicted="Effective")
not_changed_inconclusive<-inconclusive%>%filter(min_difference>0.15)

final_data2<-rbind(not_inconclusive,changed_inconclusive,not_changed_inconclusive)



final_data2<-final_data2%>%mutate(weight=0)
num_studies2<-final_data2%>%count()

for(x in 1:num_studies2$n){
  if((final_data2$studytype1[x] == "meta-analysis") & (final_data2$predicted[x]=="Effective")){final_data2$weight[x]=8}
  if((final_data2$studytype1[x]== "meta-analysis") & (final_data2$predicted[x]=="Detrimental")){final_data2$weight[x]=-8}
  if((final_data2$studytype1[x]== "meta-analysis") & (final_data2$predicted[x]=="Inconclusive")){final_data2$weight[x]=-4}
  
  if((final_data2$studytype1[x]== "clinical trial") & (final_data2$predicted[x]=="Effective")){final_data2$weight[x]=7}
  if((final_data2$studytype1[x]== "clinical trial") & (final_data2$predicted[x]=="Detrimental")){final_data2$weight[x]=-7}
  if((final_data2$studytype1[x]== "clinical trial") & (final_data2$predicted[x]=="Inconclusive")){final_data2$weight[x]=-3.5}
  
  if((final_data2$studytype1[x]== "clinical observational") & (final_data2$predicted[x]=="Effective")){final_data2$weight[x]=6}
  if((final_data2$studytype1[x]== "clinical observational") & (final_data2$predicted[x]=="Detrimental")){final_data2$weight[x]=-6}
  if((final_data2$studytype1[x]== "clinical observational") & (final_data2$predicted[x]=="Inconclusive")){final_data2$weight[x]=-3}
  
  if((final_data2$studytype1[x]== "meta-synthesis") & (final_data2$predicted[x]=="Effective")){final_data2$weight[x]=5}
  if((final_data2$studytype1[x]== "meta-synthesis") & (final_data2$predicted[x]=="Detrimental")){final_data2$weight[x]=-5}
  if((final_data2$studytype1[x]== "meta-synthesis") & (final_data2$predicted[x]=="Inconclusive")){final_data2$weight[x]=-2.5}

  if((final_data2$studytype1[x]== "clinical case report") & (final_data2$predicted[x]=="Effective")){final_data2$weight[x]=4}
  if((final_data2$studytype1[x]== "clinical case report") & (final_data2$predicted[x]=="Detrimental")){final_data2$weight[x]=-4}
  if((final_data2$studytype1[x]== "clinical case report") & (final_data2$predicted[x]=="Inconclusive")){final_data2$weight[x]=-2} 
  
  if((final_data2$studytype1[x]== "preclinical in vivo") & (final_data2$predicted[x]=="Effective")){final_data2$weight[x]=3}
  if((final_data2$studytype1[x]== "preclinical in vivo") & (final_data2$predicted[x]=="Detrimental")){final_data2$weight[x]=-3}
  if((final_data2$studytype1[x]== "preclinical in vivo") & (final_data2$predicted[x]=="Inconclusive")){final_data2$weight[x]=-1.5}
  
  if((final_data2$studytype1[x]== "preclinical in vitro") & (final_data2$predicted[x]=="Effective")){final_data2$weight[x]=2}
  if((final_data2$studytype1[x]== "preclinical in vitro") & (final_data2$predicted[x]=="Detrimental")){final_data2$weight[x]=-2}
  if((final_data2$studytype1[x]== "preclinical in vitro") & (final_data2$predicted[x]=="Inconclusive")){final_data2$weight[x]=-1}
}

scores2<-final_data2%>%group_by(drug_x,disease_cleaned)%>%
        summarize(score=sum(weight),count=n())%>%
        arrange(desc(score))
scores2

#low count exploration

scores_count<-scores2%>%mutate(score_count=score/count)%>%
              arrange(desc(score_count))
scores_count
```
```{r}
plot<-ggplot(data=scores_count)+geom_point(aes(x=count, y=score_count))
plot


```

